---
title: "Stat 159 Project 2 Presentation"
author: "Junyu Wang, Nichole Rethmeier"
date: "November 2, 2016"
output: 
  ioslides_presentation:
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

In this project, we applied model selection methods on different regression models introduced in Chapter 6: Linear Model Selection and Regularization (from **An Introduction to Statistical Learning** by James et al). The 5 regression models we practiced are _Ordinary Least Squares, Ridge Regression, Lasso Regression, Principle Components Regression and Partial Least Square Regression_. In order to find the best parameter (or coefficients) for those regression models, we splitted our datasets into train and test groups, performed 10-fold cross validation on each model respectively, and in the end chose the best model with the least cross-validation error.

## Data

The [Credit data](http://www-bcf.usc.edu/~gareth/ISL/Credit.csv) set we will be using comes from **An Introduction to Statistical Learning**, contains the following variables: Income, Limit, Rating, Cards, Age, Education, Gender, Student, Married, Ethnicity and Balance. The Credit data set records Balance, the average credit card debt for an individual, according to quantitative predictors such as Income, Limit, etc. and qualitative predictors such as Student, Married, etc. Cards represents the number of credit cards a person has and Income is represented in the thousands of dollars. We investigate 400 individuals in this analysis.

## Prediction Goal

Our prediction goal is to find the best model that fits the data. And using this model to gain insights about how different factors can affect credit balance.

## Applied Methods

- Ordinary Least Square Method
    - Simple Linear Regression
- Shrinkage Method
    - Ridge Regression
    - Lasso Regression
- Dimension Reduction Method
    - Principal Component Regression (PCR)
    - Partial Least Square Regression (PLSR)

## Analysis -- Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Analysis --  Slide with Plot

```{r pressure}
plot(pressure)
```

## Conclusion

Here is the conclusion
