
# Results

##Exploratory Data Analysis Results

```{r, echo=FALSE}
library(Matrix)
load("../data/regressions/ridge-models.RData")
load("../data/regressions/lasso-models.RData")
load("../data/regressions/plsr-models.RData")
load("../data/regressions/pcr-models.RData")
load("../data/eda-output.RData")
ridgesum <- summary(ridge_official_coef)
lassosum <- summary(lasso_official_coef)
ridge_coefficients <- data.frame(Ridge = ridgesum$x)
lasso_coefficients <- data.frame(Lasso  = lassosum$x)
all_coefficients <- data.frame(plsr_official_coef, pcr_official_coef, ridge_coefficients, lasso_coefficients)
colnames(all_coefficients) <- c("PLSR", "PCR", "Ridge", "Lasso")
```

The results of our exploratory data analysis showed that certain variables in our study were much more important when examining the `Balance` of credit card debt. Two of the variables were highly correlated with the `Balance`, a person's credit `Limit` and credit `Rating`. This is understandable the higher a person's credit rating, generally the higher their credit limit is as well which means they are at liberty to spend more leading to higher balances. Below is the correlation matrix for all of the numeric variables: 

```{r, results='asis', echo=FALSE}
library(xtable)
print(xtable(correlation_matrix, caption=c("Correlation Matrix")), comment=F)
```

##Regression Results 

After applying our regressions to our full data sets, we received the following coefficients for our various regression models. While the coefficients obviously vary across the different models, we can notice trends comparing all together. Across all regression models, the predictor that had the largest affect on Balance was `Limit`. Again across the PLSR, PCR, ridge and lasso regressions we can note that the estimates for ${\hat\beta_4, \hat\beta_5, \hat\beta_6}$ corresponding to `Age`, `Education`, and `Gender` respectively are all nearly 0. This indicates that these variables do not largely factor in to the `Balance` of a person's credit card debt. As credit card company's cannot discriminate in their providing of credit to people of different genders, ages, or with different amounts of education it makes sense that these estimates are lower.

```{r, results='asis', echo=FALSE}
print(xtable(round(all_coefficients,2), caption=c("Information about Regression Coefficients")), comment=F)
```


